Don't Mind the Gap: Bridging Network-wide Objectives and Device-level Configurations

### 2016-06-19 paper/v2/paper-edited.tex
edit section 5

% Formally define automata
%The RIR policy is an ordered sequence of regular expressions $r_1 \Prefer \dots \Prefer r_j$.
While paths in RIR policy describe the direction traffic flow through the network, to implement the policy with BGP we are concerned about the way control-plane information is disseminated, i.e., route announcements flowing in the opposite direction. To capture this idea, for each regular expression $r_i$ in RIR policy, we construct a deterministic finite state machine on the reversed regular expression. Each automaton for the reversed regular expression of $r_i$ is a tuple ($\Sigma, Q_i, F_i, q_{0_i}, \sigma_i$). The alphabet $\Sigma$ consists of all abstraction locations (i.e., routers or ASes), $Q_i$ is the set of states for automaton i, $F_i$ is the set of final states, $q_{0_i}$ is the initial state, and $\sigma_i \colon Q_i \times \Sigma \rightarrow Q_i$ is the state transition function.
%
% Formally define topology
The topology is represented as a graph ($V, E$), which consists of a set of vertices $V$, and a set of directed edges $E \colon V \times V$.
%
% Formally define product graph
The combined PGIR is a tuple ($V'$, $E'$, $s$, $e$, $P$) with
vertices $V' \colon V \times Q_1 \times \dots \times Q_j$,
edges $E' \colon V' \times V'$,
a unique starting vertex $s$,
a unique ending vertex $e$,
and a preference function $P \colon V' \rightarrow 2^{\set{1, \dots, j}}$ , which maps nodes in the product graph to a set of preference ranks.
For a PGIR vertex $v_1 = (v, \dots) \in V'$, we say $loc(v_1) = v$, which indicates that the corresponding topology location for $v_1$ is $v$.

%
We say that a PGIR node $x_1$ is a \emph{shadow} of topology location $x$ when $x_1 \in V'$ and $loc(x_1) = x$.
%
%Throughout the remainder of the paper, we will use metavariables without subscripts such as $x$ to represent topology locations, and we will use similar metavariables with subscripts such as $x_1$ or $x_2$ to represent shadows of $x$ (i.e., PGIR nodes for location $x$).

@@

We use graph dominators~\cite{tarjan-dominance} as a cheap approximation for removing many nodes and edges in the PGIR that are never on any \emph{simple} (loop free) path between the start and end nodes.
In the PGIR, a node $x$ dominates a node $y$ if $x$ appears on every path leading from the start node to $y$. Similarly, a node $x$ post-dominates a node $y$ in the PGIR if $x$ appears on every path from $y$ to the end node.
We can safely remove nodes and edges in the PGIR when any of the following conditions hold, where $x$, $x'$ and $y$, $y'$ refer to PGIR nodes that are shadows of the same topology location.
\begin{itemize}
\setlength{\itemsep}{1pt}
\setlength{\parskip}{0pt}
\setlength{\parsep}{0pt}
\item Remove $x$ if it is not reachable from the start node
\item Remove $x$ if it can not reach the end node
\item Remove $x$ if it is (post-)dominated by some $x'$
\item Remove edge ($x$, $y$) if $y$ is post-dominated by some $x'$
\item Remove edge ($x$, $y$) if $x$ is dominated by some $y'$
\end{itemize}
For example, node $(W,1,1)$ in Figure~\ref{fig:example-compilation} is removed since it is never on a simple path to the end node because the path must always go through node $(W,-,4)$. That is, node $(W,-,4)$ post-dominates node $(W,1,1)$ and both shadow topology location $W$. 
Similarly, the edge from $(C,3,2)$ to $(D,-,2)$ is removed since node $(C,-,2)$ post-dominates $(D,-,2)$ and shadows the same topology location $C$. 
We repeatedly apply the minimizations above until no further minimization is possible.
In the example of Figure~\ref{fig:example-compilation}, colored nodes and dashed edges are removed after minimization.

@@

The problem of determining local preferences for each router is closely reflected in the structure of the PGIR. In particular, whenever a given router appears in multiple different contexts in the PGIR (i.e., as multiple PGIR nodes that shadow each other), then the compiler must decide which context to prefer.
%
In our compilation example from Figure~\ref{fig:example-compilation}, the topology node $C$ can receive an advertisement from $E$ in context $(C,-,2)$ or it can receive an advertisement from its neighbor $D$ in context $(C,3,2)$.
%
The compiler's task is to determine a \emph{total ordering} of contexts for each router in the topology, which reflects the relative preference of advertisements received in each context. For example, if the compiler determines that, for router $C$, context $(C,3,2)$ is preferred to context $(C,-,2)$, written as $(C,3,2) <_{pref} (C,-,2)$, then this means that $C$ will prefer advertisements heard from $D$ (with tag $(2,2)$) over $E$ (with tag $(-,2)$). Section~\ref{sec:abgp} discusses our encoding of preferences and filters in more detail.

@@

A natural question then is, how should the compiler order PGIR contexts for each router in a way that is both \emph{sound} and \emph{complete} under all possible failures? To this end, we introduce the notion of \emph{regret-free} preferences, motivated by the observations from the example in Figure~\ref{fig:unimplementable}. Router $r$ has a \emph{regret-free} preference for a set of advertisements $X$ over $Y$ if, whevever $r$ selects an advertisement from $X$ over another from $Y$, there is always some policy-compliant path to each final (advertisement) destination $d$ that is at least as good $(\leq_{rank})$ as any possible path to $d$ if it had accepted an advertisement from $Y$ instead. In other words, the preference of $X$ over $Y$ at $r$ is regret-free if we are never worse off by choosing from $X$.
%
In the compilation example from Figure~\ref{fig:example-compilation}, the choice for $C$ to prefer context $(C,3,2)$ to $(C,-,2)$ is regret-free, since there will always be at least as good a path to destination $W$ regardless of any failures that might occur in the network.

@@

\begin{figure}[t!]
  \centering
  \includegraphics[width=.5\columnwidth]{figures/unimplementable}
  \caption{A network where the policy $(\CD{A} \cdot \CD{B} \cdot \CD{D} \cdot \CD{E} \cdot \CD{G}) \Prefer (\CD{A} \cdot \CD{C} \cdot \CD{D} \cdot \CD{F} \cdot \CD{G})$ is unimplementable in BGP under arbitrary failures.}
  \label{fig:unimplementable}
  %\vspace{-1em}
\end{figure}

@@

Searching for \emph{regret-free} preferences in general is hard, and enumerating all possible combinations of failures and preference orderings is intractable. We thus adopt a conservative analysis that we found to be both effective and efficient in practice. The idea is to $i)$ search for regret-free preferences by comparing the set of paths available after accepting advertisments in two different PGIR contexts $N_1$ and $N_2$, and $ii)$ refine the comparison when necessary by considering where the announcements must have traversed before arriving at $N_1$ or $N_2$.

\begin{algorithm}[t!]
\caption{Inferring regret-free preferences}
\label{alg:failures}
\begin{algorithmic}[1]
\Procedure{Regret-free($G$, $N_1$, $N_2$)}{}
  \If {$loc(N_1) \neq loc(N_2)$} \Return false
  \EndIf
  \State $q \gets Queue()$
  \State $q.Enqueue (N_1, N_2)$
  \While {$!q.Empty()$}
    \State $(n_1,n_2) \gets q.Dequeue()$
    \If {$n_1 \nleq_{rank} n_2$}
      \Return false
    \EndIf
    \For {$x$ in adj($G$, $n_2$)}
      \If {\big($\exists y \in$ adj($G$,$n_1$), shadows $x$\big) \textbf{or} \\
            \hspace{5.2em} \big($\exists y \in$ $G$, dominates $n_1$, shadows $x$\big)}
        \If {$(x,y)$ not marked}
          \State {mark $(x,y)$ as seen}
          \State $q.Enqueue(x,y)$
        \EndIf
      \Else { \Return $false$}
      \EndIf
    \EndFor
  \EndWhile
  \Return true
  \EndProcedure
\end{algorithmic}
\end{algorithm}

Algorithm~\ref{alg:failures} checks whether one context can be preferred to another ($N_1 <_{pref} N_2$). The algorithm walks from nodes $N_1$ and $N_2$ and ensures that for every \textit{step} $N_2$ can take to some new topology location, $N_1$ can, at the very least, also take a step to an equivalent topology location. When there is no such equivalent step, the algorithm attempts to take into account where the advertisement must have already traversed. In particular, it checks to see if there is an equivalent dominator (a topology location the advertisement must have passed through earlier) and walks from this new node instead.
At each step, it requires that current node reachable from $N_1$ has a path rank that is at least as good as that of the current node reachable from $N_2$ ($n_1 \leq_{rank} n_2$). The intuition here is that if $n_1 \nleq_{rank} n_2$, then we can likely fail every edge in the topology except for the path that leads to the current $n_1$ and $n_2$, thereby generating in a counter example.
Algorithm~\ref{alg:failures} terminates since the number of related states $(x,y)$ that can be explored is finite.

For each router in the topology, local preferences are now obtained by sorting the corresponding PGIR contexts according to the $(<_{pref})$ relation. If two nodes are incomparable, then the compiler rejects the policy as unimplementable.

@@

The checks for failure safety described above overlooked one critical point: A better (lower rank) path might not be available due to loops that are rejected by BGP. To avoid this situation, the compiler checks that, any time some context $N_1$ appears ``above'' (i.e., can reach) $N_2$ in the PGIR, then $N_1$ must be strictly preferred to $N_2$. The idea is that anytime $N_2$ is not usable because the advertisement already went through $N_1$, it ultimately does not matter since $N$ has a regret-free preference for $N_1$ anyway.

@@

%\begin{figure}[t!]
%\begin{lstlisting}[frame=single, mathescape=true]
%$\Router{A}$
%  Match $\PEER=C$
%    Export $\MED \leftarrow 80, \PEER \leftarrow W$
%$\Router{B}$
%  Match $\PEER = C$
%    Export $\COMM \leftarrow \text{noexport}, \MED \leftarrow 81, \PEER \leftarrow W$
%$\Router{C}$
%  Match$[LP=99]$ $\PEER = E $
%    Export $\PEER \leftarrow B$
%  Match $\PEER = D$
%    Export $\PEER \leftarrow A,B$
%$\Router{D}$
%  Match $\REGEX{X + Y}$
%    Export $\PEER \leftarrow C$
%$\Router{E}$
%  Match $\REGEX{Z}$
%    Export $\PEER \leftarrow C$
%\end{lstlisting}
%\label{fig:config-min}
%\caption{Abstract BGP minimized configurations}
%\end{figure}

>>>>

% Formally define automata
%The RIR policy is an ordered sequence of regular expressions $r_1 \Prefer \dots \Prefer r_j$.
While paths in RIR policy describe the direction traffic flow through the network, to implement the policy with BGP we are concerned about the way control-plane information is disseminated, i.e., route announcements flowing in the opposite direction. To capture this idea, for each regular expression $r_i$ in RIR policy, we construct a deterministic finite state machine on the reversed regular expression. Each automaton for the reversed regular expression of $r_i$ is a tuple ($\Sigma, Q_i, F_i, q_{0_i}, \sigma_i$). The alphabet $\Sigma$ consists of all abstraction locations (i.e., routers or ASes), $Q_i$ is the set of states for automaton i, $F_i$ is the set of final states, $q_{0_i}$ is the initial state, and $\sigma_i \colon Q_i \times \Sigma \rightarrow Q_i$ is the state transition function.
%
% Formally define topology
The topology is represented as a graph ($V, E$), which consists of a set of vertices $V$, and a set of directed edges $E \colon V \times V$.
%
% Formally define product graph
The combined PGIR is a tuple ($V'$, $E'$, $s$, $e$, $P$) with
vertices $V' \colon V \times Q_1 \times \dots \times Q_j$,
edges $E' \colon V' \times V'$,
a unique starting vertex $s$,
a unique ending vertex $e$,
and a preference function $P \colon V' \rightarrow 2^{\set{1, \dots, j}}$ , which maps nodes in the product graph to a set of preference ranks.
For a PGIR vertex $v' = (v, \dots) \in V'$, we say $loc(v') = v$ to indicate that the topology location for $v'$ is $v$.

%
We say that a PGIR node $x_1$ is a \emph{shadow} of topology location $x$ when $x_1 \in V'$ and $loc(x_1) = x$. As a shorthand, we write $x_1 \approx x_2$ when $x_1$ and $x_2$ are both shadows for the same topology location.
%
%Throughout the remainder of the paper, we will use metavariables without subscripts such as $x$ to represent topology locations, and we will use similar metavariables with subscripts such as $x_1$ or $x_2$ to represent shadows of $x$ (i.e., PGIR nodes for location $x$).

@@

We use graph dominators~\cite{tarjan-dominance} as a cheap approximation for removing many nodes and edges in the PGIR that are never on any \emph{simple} (loop free) path between the start and end nodes.
In the PGIR, a node $x$ dominates a node $y$ if $x$ appears on every path leading from the start node to $y$. Similarly, a node $x$ post-dominates a node $y$ in the PGIR if $x$ appears on every path from $y$ to the end node.
We can safely remove nodes and edges in the PGIR when any of the following conditions hold, where $x$, $x'$ and $y$, $y'$ refer to PGIR nodes that are shadows of the same topology location.
\begin{itemize}
\setlength{\itemsep}{1pt}
\setlength{\parskip}{0pt}
\setlength{\parsep}{0pt}
\item Remove $x$ if it is not reachable from the start node
\item Remove $x$ if it can not reach the end node
\item Remove $x$ if it is (post-)dominated by some $x'$
\item Remove edge ($x$, $y$) if $y$ is post-dominated by some $x'$
\item Remove edge ($x$, $y$) if $x$ is dominated by some $y'$
\end{itemize}
For example, node $(W,1,1)$ in Figure~\ref{fig:example-compilation} is removed since it is never on a simple path to the end node because the path must always go through node $(W,-,4)$. That is, node $(W,-,4)$ post-dominates node $(W,1,1)$ and both are shadows of topology location $W$. 
Similarly, the edge from $(C,3,2)$ to $(D,-,2)$ is removed since node $(C,-,2)$ post-dominates $(D,-,2)$ and is a shadow of the same topology location $C$. 
We repeatedly apply the minimizations above until no further minimization is possible.
In the example from Figure~\ref{fig:example-compilation}, colored nodes and dashed edges show edges and nodes removed after minimization.

@@

The problem of determining local preferences for each router is closely reflected in the structure of the PGIR. In particular, whenever a given router appears in multiple different contexts in the PGIR (i.e., as multiple PGIR nodes that are shadows of the router), then the compiler must decide which shadow to prefer.
%
In our compilation example from Figure~\ref{fig:example-compilation}, the topology node $C$ can receive an advertisement from $E$ in shadow $(C,-,2)$ or it can receive an advertisement from its neighbor $D$ in shadow $(C,3,2)$.
%
The compiler's task is to determine a \emph{total ordering} of shadow nodes for each router in the topology, which reflects the relative preference of advertisements received in each shadow. For example, if the compiler determines that, for router $C$, shadow $(C,3,2)$ can be locally preferred to shadow $(C,-,2)$, written as $(C,3,2) \leq_{lp} (C,-,2)$, then this means that $C$ will prefer advertisements heard from $D$ (with tag $(2,2)$) over $E$ (with tag $(-,2)$). Section~\ref{sec:abgp} discusses our encoding of preferences and filters in more detail.

@@

A natural question then is, how should the compiler order PGIR shadows ($\leq_{lp}$) for each router in a way that is both \emph{sound} and \emph{complete} under all possible failures? To this end, we introduce the notion of \emph{regret-free} preferences, motivated by the observations from the example in Figure~\ref{fig:unimplementable}. We say a router $r$ has a \emph{regret-free} preference for a set of advertisements $X$ over $Y$ if, whevever $r$ selects an advertisement from $X$ over another from $Y$, there is always some policy-compliant path to each final (advertisement) destination $d$ that is at least as good $(\leq_{rank})$ as any possible path to $d$ if it had accepted an advertisement from $Y$ instead. In other words, the preference of $X$ over $Y$ at $r$ is regret-free if we are never worse off by choosing an advertisement from $X$ when available. The notion of regret-free preferences can be lifted to nodes in the PGIR by considering the set of advertisments available to each node.

In the compilation example from Figure~\ref{fig:example-compilation}, the choice for $C$ to prefer shadow $(C,3,2)$ to $(C,-,2)$ is regret-free, since there will always be at least as good a path to destination $W$ regardless of any failures that might occur in the network. For example, if the $C$--$A$ link fails, then there is still a backup path from $(C,3,2)$ to $(W,-,4)$ that is just as good as any path from $(C,-,2)$. Likewise, any combination of failures to disconnect $(C,3,2)$ from $(W,-,4)$ would also disconnect $(C,-,2)$ from $(W,-,4)$.

@@

\begin{figure}[t!]
  \centering
  \includegraphics[width=.8\columnwidth]{figures/unimplementable}
  \caption{A network where the policy $(\CD{A} \cdot \CD{B} \cdot \CD{D} \cdot \CD{E} \cdot \CD{G}) \Prefer (\CD{A} \cdot \CD{C} \cdot \CD{D} \cdot \CD{F} \cdot \CD{G})$ is unimplementable in BGP under arbitrary failures.}
  \label{fig:unimplementable}
  %\vspace{-1em}
\end{figure}

@@

Searching precisely for \emph{regret-free} preferences in general is hard, and clearly enumerating all possible combinations of failures and preference orderings is intractable. We thus adopt a conservative analysis that we found to be both effective and efficient in practice. The idea is to $i)$ search for regret-free preferences by comparing the set of paths available after accepting advertisments in two different PGIR shadows $N_1$ and $N_2$ of topology location $N$, and $ii)$ refine the comparison when necessary by considering where the announcements must have traversed before arriving at $N_1$ or $N_2$.

\begin{algorithm}[t!]
\caption{Inferring regret-free preferences}
\label{alg:failures}
\begin{algorithmic}[1]
\Procedure{Regret-free($G$, $N_1$, $N_2$)}{}
  \If {$loc(N_1) \neq loc(N_2)$} \Return false
  \EndIf
  \State $q \gets Queue()$
  \State $q.Enqueue (N_1, N_2)$
  \While {$!q.Empty()$}
    \State $(n_1,n_2) \gets q.Dequeue()$
    \If {$n_1 \nleq_{rank} n_2$}
      \Return false
    \EndIf
    \For {$n_2'$ in adj($G$, $n_2$)}
      \If {\big($\exists n_1' \in \textup{adj}(G,n_1), n_1' \approx n_2'$\big) \textbf{or} \\
            \hspace{5.2em} 
            \big($\exists n_1' \in$ $G$, dominates $n_1$, $n_1' \approx n_1'$\big)}
        \If {$(x,y)$ not marked}
          \State {mark $(x,y)$ as seen}
          \State $q.Enqueue(x,y)$
        \EndIf
      \Else { \Return $false$}
      \EndIf
    \EndFor
  \EndWhile
  \Return true
  \EndProcedure
\end{algorithmic}
\end{algorithm}

Algorithm~\ref{alg:failures} checks whether one shadow can be preferred to another ($N_1 \leq_{lp} N_2$). The algorithm walks from nodes $N_1$ and $N_2$ and ensures that for every \textit{step} $N_2$ can take to some new topology location, $N_1$ can, at the very least, also take a step to an equivalent topology location $(\approx)$. When there is no such equivalent step, the algorithm attempts to take into account where the advertisement must have already traversed. In particular, it checks to see if there is an equivalent dominator (a topology location the advertisement must have passed through earlier) and walks from this new node instead.
At each step, it requires that current node reachable from $N_1$ has a path rank that is at least as good as that of the current node reachable from $N_2$ ($n_1 \leq_{rank} n_2$). The intuition here is that if $n_1 \nleq_{rank} n_2$, then we can very likely fail every edge in the topology except for the path that leads to the current $n_1$ and $n_2$, thereby generating in a counter example.
Algorithm~\ref{alg:failures} terminates since the number of related states $(n_1,n_2)$ that can be explored is finite.

For each router in the topology, local preferences are now obtained by sorting the corresponding PGIR contexts according to the $(\leq_{lp})$ relation determined by the \textup{Regret-Free}. If two nodes are incomparable, then the compiler rejects the policy as unimplementable.

@@

The checks for failure safety described above overlooked one critical point: A better (lower rank) path might not be available due to loops that are rejected by BGP. 
%
For example, in Figure~\ref{fig:example-compilation}, suppose we did not remove node $(D,-,2)$ during minimization. In this case, $(C,3,2)$ might think that $(D,-,2)$ is available for use along a path to $W$ even though the advertisement might have already gone through $(D,2,2)$ prior to reaching $(C,3,2)$. Since BGP would reject any path using $D$ twice, the PGIR node $(D,-,2)$ may be unusable for $(C,3,2)$ despite the analysis not taking this information into account.

To avoid this situation, the compiler checks that, any time we have two shadows $N_1$ and $N_2$ for some topology location $N$, where $N_1$ appears ``above'' (i.e., can reach) $N_2$ in the PGIR, then $N_1$ must be strictly preferred to $N_2$ (that is, $N_1 <_{lp} N_2$). The idea is that anytime $N_2$ is not usable because the advertisement already went through $N_1$, it ultimately does not matter since $N$ has a regret-free preference for $N_1$ anyway.
%
For example, $(D,2,2)$ appears ``above'' $(D,-,2)$, so we require that $(D,2,2) <_{lp} (D,-,2)$. In this case, the compiler determines that this is safe according to Algorithm~\ref{alg:failures}, so router $D$ can simply prefer advertisements from $Y$ and $X$ over advertisements from $C$.

@@
